---
title: "Lab 1"
author: "Laya Buchanan"
date: "1/14/2021"
output: github_document
---

Lab goal is to create a county level map of Quality of Life Index Ranking from the Robert Wood Johnson Foundation (obtained from here: https://www.countyhealthrankings.org/)

```{r}
library(tidyverse)
library(data.table)
library(tidycensus)
library(ggplot2)
library(RColorBrewer)
census_api_key("aa512886c5449a582d837da8d3a07af66a043fe5")

census_data = load_variables(2018, "acs5", cache=T)
fwrite(census_data, "census_variables.csv")

vars = c(tpop = 'P001001',
          medage = 'P013001',
          wpop = 'P003002',
          bpop = 'P003003',
          apop = 'P003005',
          hpop = 'P004003')

NY_df = get_decennial(state = "ny", 
                       geography = "county",
                       variables = vars,
                       geometry = T,
                       output = "wide")

NY_df = 
  NY_df %>% 
	  mutate(
		  wpct = (wpop/tpop) *100,
		  bpct = (bpop/tpop) *100,
		  apct = (apop/tpop) *100,
		  hpct = (hpop/tpop) *100
	)
```

Compute Simpson's Diversity Index **D = ((SUM n(n-1))/N(N-1))** Higher value indicates higher diversity
```{r}
NY_df = 
  NY_df %>% 
  mutate(
    race_div = 1 - (((wpop*(wpop-1))+(bpop*(bpop-1))+(hpop*(hpop-1))+(apop*(apop-1)))/(tpop*(tpop-1)))
  )

plot(NY_df["race_div"])


rwj = fread("rwj_rank.csv", 
             stringsAsFactors = F, 
             data.table = F, 
             colClasses=list(character=c("FIPS")))

ny_rwj = subset(rwj, State == "New York")


ny_rwj_df =
  merge(NY_df, 
        ny_rwj,
        by.x = "GEOID",
        by.y = "FIPS")

summary(ny_rwj_df)
summary(ny_rwj_df$QL.Rank)

ny_rwj_df = 
  ny_rwj_df %>% 
  mutate(QL.Rank = as.numeric(QL.Rank))

summary(ny_rwj_df$QL.Rank)


```
### Visualize

#####use ggplot to visualize data in map

map1
```{r}
map1 = ny_rwj_df %>% 
  ggplot() +
  geom_sf(aes(fill = cut_number(QL.Rank, 5)))
```

map 2
```{r}
map2 =
  ny_rwj_df %>% 
  ggplot(aes(fill = QL.Rank)) +
  geom_sf() +
  scale_fill_continuous(low = "#34E8EB", high = "#3D34EB") +
  #scale_fill_continuous(low = "lightgreen", high = "darkgreen") +
  ggtitle("County Level Quality of Life Rank") +
  theme(line = element_blank(),                          
        axis.text=element_blank(),                       
        axis.title=element_blank(),                      
        panel.background = element_blank())

summary(ny_rwj_df$QL.Rank)
ny_rwj_df = 
ny_rwj_df %>% 
  mutate(
    QL.Rank = (QL.Rank - 63) * -1
  )
```
map 3
```{r}
map3 =
  ny_rwj_df %>% 
  ggplot() +
  geom_sf(aes(fill=QL.Rank)) +
  scale_fill_continuous(low = "lightgreen", high = "darkgreen") +
  ggtitle("County Level Quality of Life Rank") +
  theme(axis.text=element_text(size=8),                       
        axis.title=element_text(size=8,face="bold"),
        plot.title = element_text(hjust = 0.5))  
```

map 4
```{r}
map4 =
  ny_rwj_df %>% 
  ggplot() +
  geom_sf(aes(fill=QL.Rank)) +
  scale_fill_continuous(low = "lightgreen", high = "darkgreen") +
  ggtitle("County Level Quality of Life Rank") +
  theme(axis.text=element_text(size=8),                       
        axis.title=element_text(size=8,face="bold"),
        plot.title = element_text(face="bold",size=12,hjust = 0.5)) 
map4
```

map 5
```{r}
map5 =
  ny_rwj_df %>% 
  ggplot() +
  geom_sf(aes(fill=QL.Rank)) +
  scale_fill_continuous(low = "lightgreen", high = "darkgreen") +
  ggtitle("County Level Quality of Life Rank") +
  theme(axis.text=element_text(size=8),                       
        axis.title=element_text(size=8,face="bold"),
        plot.title = element_text(face="bold",size=12,hjust = 0.5),
        axis.ticks = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_blank(),
        legend.position = c(0.9,0.5)) 
```

map 6
```{r}
map6 = ny_rwj_df %>% 
  ggplot() +
  geom_sf(aes(fill=QL.Rank)) +
  scale_fill_continuous(low = "lightgreen", high = "darkgreen",
                        breaks = c(12, 24, 36, 48, 60),
                        name = "Quality of Life",
                        labels = c("low",
                                   "",
                                   "",
                                   "",
                                   "high")) +
  ggtitle("County Level Quality of Life Rank") +
  theme(axis.text=element_text(size=8),                       
        axis.title=element_text(size=8,face="bold"),
        plot.title = element_text(face="bold",size=12,hjust = 0.5),
        axis.ticks = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_blank(),
        legend.position = c(0.9,0.5)) 
```

map 7
```{r}
map7 = 
  ny_rwj_df %>% 
  ggplot() +
  geom_sf(aes(fill=QL.Rank)) +
  scale_fill_continuous(low = "lightgreen", high = "darkgreen",
                        breaks = c(3, 12, 24, 36, 48, 60),
                        name = "Quality\nof\nLife",
                        labels = c("low",
                                   "",
                                   "",
                                   "",
                                   "",
                                   "high")) +
  ggtitle("County Level Quality of Life Rank") +
  theme(axis.text=element_text(size=8),                       
        axis.title=element_text(size=8,face="bold"),
        plot.title = element_text(face="bold",size=12,hjust = 0.5),
        axis.ticks = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_blank(),
        legend.position = c(0.9,0.5),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8)) 

names(ny_rwj_df)
summary(ny_rwj_df$race_div)
```

map 8
```{r}
map8 = 
  ny_rwj_df %>% 
  ggplot() +
  geom_sf(aes(fill = race_div)) +
  scale_fill_continuous(low = "lightgreen", high = "darkgreen",
                        breaks = c(0.05, 0.1, 0.16, 0.31, 0.65),
                        name = "Racial\nDiversity",
                        labels = c("low",
                                   "",
                                   "",
                                   "",
                                   "high")) +
  ggtitle("County Level Racial Diversity") +
  theme(axis.text=element_text(size=8),                       
        axis.title=element_text(size=8,face="bold"),
        plot.title = element_text(face="bold",size=12,hjust = 0.5),
        axis.ticks = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_blank(),
        legend.position = c(0.9,0.5),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8))
```

plot 1

```{r}
ny_rwj_df %>% 
  mutate(
    race_div_bin = ifelse(race_div > mean(race_div),
                          "High Diversity",
                          "Low Diversity"),
    Qual_Life_bin = ifelse(QL.Rank > mean(QL.Rank),
                           "High Quality of Life",
                           "Low Quality of Life")
  )

plot1 = 
  ny_rwj_df %>% 
  ggplot(aes(x = race_div_bin, 
             y = QL.Rank, 
             group = race_div_bin,
             color = race_div_bin,
             fill = race_div_bin)) + 
  geom_boxplot() + 
  labs(title="Boxplots", 
       subtitle= "Quality of Life by Racial Diversity",
       y="Quality of Life", 
       x="Racial Diversity") +
  scale_colour_brewer(palette = "Set3") +
  theme(legend.position = "none")
```

plot2
```{r}
plot2 = 
  ny_rwj_df %>% 
  ggplot(aes(x = Qual_Life_bin,
             y = race_div,
             group = Qual_Life_bin,
             color = Qual_Life_bin,
             fill = Qual_Life_bin)) + 
  geom_boxplot() + 
  labs(title="Boxplots", 
       subtitle= "Racial Diversity by Quality of Life",
       y="Racial Diversity", 
       x="Quality of Life") +
  scale_colour_brewer(palette = "Set3") +
  theme(legend.position = "none")
```

full plot
```{r}
library(ggpubr)

full_plot = ggarrange(map7,
                       map8,
                       plot1,
                       plot2,
                       ncol = 2,
                       nrow = 2)

ggsave("full_plot.pdf")
```


Please submit the result of the lab script in 
fulfillment of your assignment 1 requirement





